---
title: "Report: `r params$project`"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    toc_depth: 3
params:
  project: test-gene-level
  date: 2022-05-27
  version: v0.4.2
  regenie_merged: ../output/test-gene-based-test/results/Y1.regenie.gz
  regenie_filename: test-gene-level.Y1.regenie.all.gz
  phenotype: Y1
  covariates:
  phenotype_file: ../output/test-gene-based-test/validated_input/phenotype_bin.validated.txt
  regenie_step1_log: ../output/test-gene-based-test/logs/test-gene-based-test.step1.log
  regenie_step2_log: ../output/test-gene-based-test/logs/test-gene-based-test.step2.log
  phenotype_log: ../output/test-gene-based-test/logs/phenotype_bin.validated.log
  covariate_log: ../output/test-gene-based-test/logs/covariates.validated.log
  plot_ylimit: 0
  mask_file: ./example.masks
  manhattan_annotation_enabled: true
  annotation_min_log10p: 1
  r_functions: ./functions.R
  rmd_pheno_stats: ./child_phenostatistics.Rmd
  rmd_valdiation_logs: ./child_validationlogs.Rmd
---

```{r setup, include=FALSE}
library(qqman)
library(DT)
library(data.table)
library(R.utils)
library(ggplot2)
library(skimr)
library(kableExtra)
library(dplyr)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, error = TRUE, cache=FALSE)
source(params$r_functions)
```


## Project Summary

| Parameter        | Value                       |
|------------------|-----------------------------|
| Project          | `r params$project`          |
| Pipeline Version | `r params$version`          |
| Date             | `r params$date`             |
| Phenotype File   | `r params$phenotype_file`   |
| Phenotype        | `r params$phenotype`        |
| Covariates       | `r params$covariates`       |
| Mask File        | `r params$mask_file`        |
| Regenie Output   | `r params$regenie_filename` |


## Phenotype Statistics

```{r child=params$rmd_pheno_stats}

```

## Gene-based settings

```{r}
regenieTable <- fread(params$regenie_merged, header=TRUE, tmpdir="$PWD")
regenieTable <- regenieTable %>% tidyr::separate(ID,
                                                 into=c("Set", "Mask", "AAF"),
                                                 sep="\\.",
                                                 extra="merge")

regenieTable_burdenJoint <- regenieTable[is.na(regenieTable$Mask) == TRUE, ]
regenieTable <- regenieTable[is.na(regenieTable$Mask) == FALSE, ]
regenieTable <- regenieTable %>% tidyr::unite(.,
                                              "Settings",
                                              TEST, AAF, Mask,
                                              sep = ".",
                                              remove = FALSE)
```

### Tests

```{r}
gene_tests <- unique(regenieTable$TEST)
kable(gene_tests,
      col.names = c("Tests"))
```


### AAF cutoffs

```{r}
aaf_cutoffs <- unique(regenieTable$AAF)
kable(aaf_cutoffs,
      col.names = c("AAF cutoffs"))
```

### Masks
```{r}
masks <- read.table(params$mask_file, header=FALSE)
names(masks) <- c("Mask", "Annotations")
kable(masks)
```

### Joint tests for burden masks

```{r, results='asis'}
if(nrow(regenieTable_burdenJoint) > 0) {
test_joint <- unique(regenieTable_burdenJoint$TEST)
test_joint <- test_joint %>% tidyr::separate(TEST, into=c("Model", "Test", "Joint"), sep="-.", extra="merge")
cat("The following joint test was used for burden masks:", unique(test_joint$Joint))
} else {
    cat("No joint test was performed for burden masks and AAF thresholds.")
}

```


```{r, eval=nrow(regenieTable_burdenJoint) > 0, results='asis'}
cat('\n\n')
cat('## Joint test for burden masks and AAF thresholds')
cat('\n\n')
if(params$plot_ylimit == 0) {
  limit = NULL
  } else {
    limit = c(0, params$plot_ylimit )
    cat(paste("Please note that the Y-axis limit has been set to '", params$plot_ylimit,"'. Please use the 'plot_ylimit' parameter to adapt this behaviour."))
      }
#manhattan plot
cat('\n\n')
cat('**Manhattan Plot**')
cat('\n\n')
df_prep <- manhattan_prep(regenieTable_burdenJoint)
plot <- manhattan_plot(df_prep, ymax = params$plot_ylimit, hlines = FALSE)
print(plot)
#annotatate plot (if it is enabled)
if(!params$manhattan_annotation_enabled) {
  plot
  } else {
    plot <- plot + ggrepel::geom_text_repel(data = df_prep %>% group_by(Set) %>%
                             top_n(1, LOG10P) %>%
                               slice(1) %>%
                             filter(LOG10P > params$annotation_min_log10p),
                            aes(label=Set), color = 1, fontface = "italic")
    }
#qqplot
cat('\n\n')
cat('**QQ Plot**')
cat('\n\n')
pp <- qq_plot(df=regenieTable_burdenJoint)
print(pp)

```


```{r, results='asis', dpi=300}
for (test in unique(regenieTable$TEST)){
  regenieTable_tmp01 <- regenieTable[regenieTable$TEST == test, ]
  cat('\n\n')
  cat('## Test: ', unique(regenieTable_tmp01$TEST))
  cat('\n\n')
  for (settings in unique(regenieTable_tmp01$Settings)) {
    regenieTable_tmp02 <- regenieTable_tmp01[regenieTable_tmp01$Settings == settings, ]
    annotations <- masks[masks$Mask == unique(regenieTable_tmp02$Mask),]$Annotations
    cat('\n\n')
    cat('### AAF cutoff: ', unique(regenieTable_tmp02$AAF), ', Mask: ', unique(regenieTable_tmp02$Mask), ' (', annotations, '), (Test: ', unique(regenieTable_tmp02$TEST), ')',
        sep = "")
    cat('\n\n')
    #note for manual ylim setting (if ymax is defined)
    if(params$plot_ylimit == 0) {
      limit = NULL
      } else {
        limit = c(0, params$plot_ylimit )
        cat(paste("Please note that the Y-axis limit has been set to '", params$plot_ylimit,"'. Please use the 'plot_ylimit' parameter to adapt this behaviour."))
      }
    #manhattan plot
    cat('\n\n')
    cat('**Manhattan Plot**')
    cat('\n\n')
    df_prep <- manhattan_prep(regenieTable_tmp02)
    plot <- manhattan_plot(df_prep, ymax = params$plot_ylimit, hlines = FALSE)
    print(plot)
    #annotatate plot (if it is enabled)
    if(!params$manhattan_annotation_enabled) {
      plot
      } else {
        plot <- plot + ggrepel::geom_text_repel(data = df_prep %>% group_by(Set) %>%
                             top_n(1, LOG10P) %>%
                               slice(1) %>%
                             filter(LOG10P > params$annotation_min_log10p),
                            aes(label=Set), color = 1, fontface = "italic")
        }
    #qqplot
    cat('\n\n')
    cat('**QQ Plot**')
    cat('\n\n')
    pp <- qq_plot(df=regenieTable_tmp02)
    print(pp)
  }
}
```



## Validation and Logs

```{r child=params$rmd_valdiation_logs}

```


------------------------------------------------------------------------

<small>
This report has been created with **[nf-gwas `r params$version`](https://github.com/genepi/nf-gwas)**, a nextflow pipeline developed by [Sebastian Schönherr](mailto:sebastian.schoenherr@i-med.ac.at), [Lukas Forer](mailto:lukas.forer@i-med.ac.at) and [Johanna Schachtl-Riess](mailto:johanna.schachtl-riess@i-med.ac.at), Institute of Genetic Epidemiology, Medical University of Innsbruck.
</small>
