---
title: "Report: `r params$project`"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
    toc_depth: 3
params:
  project: test-gwas
  date: 2021-08-09
  version: v0.4.2
  regenie_merged: ../output/test-gwas/results/Y1.regenie.gz
  regenie_filename: test-gwas.Y1.regenie.all.gz
  phenotype: Y1
  covariates:
  interaction_gxe: 
  interaction_gxg:
  phenotype_file: ../output/test-gwas/validated_input/phenotype.validated.txt
  regenie_step1_log: ../output/test-gwas/logs/test-gwas-additive.step1.log
  regenie_step2_log: ../output/test-gwas/logs/test-gwas-additive.step2.log
  phenotype_log: ../output/test-gwas/logs/phenotype.validated.log
  covariate_log: NO_COV_LOG
  plot_ylimit: 0
  manhattan_annotation_enabled: true
  annotation_min_log10p: 5
  annotated_tophits_filename: ../output/test-gwas/results/tophits/Y1.regenie.filtered.annotated.txt.gz
  r_functions: ./functions.R
  rmd_pheno_stats: child_phenostatistics.Rmd
  rmd_valdiation_logs: child_validationlogs.Rmd
---

```{r setup, include=FALSE}
library(DT)
library(data.table)
library(R.utils)
library(ggplot2)
library(skimr)
library(kableExtra)
library(dplyr)
library(ramwas)
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, error = TRUE, cache=FALSE)
source(params$r_functions)
```

## Project Summary

| Parameter                | Value                       |
|--------------------------|-----------------------------|
| Project                  | `r params$project`          |
| Pipeline Version         | `r params$version`          |
| Date                     | `r params$date`             |
| Phenotype File           | `r params$phenotype_file`   |
| Phenotype                | `r params$phenotype`        |
| Covariates               | `r params$covariates`       |
| Interaction variable GxE | `r params$interaction_gxe`  |
| Interaction variable GxG | `r params$interaction_gxg`  |
| Regenie Output           | `r params$regenie_filename` |


## Phenotype Statistics

```{r child=params$rmd_pheno_stats}

```

## Results

```{r, results='asis', dpi=300}
regenieTable <- fread(params$regenie_merged, select = c("CHROM","GENPOS","LOG10P","ID", "A1FREQ", "TEST"), header=TRUE, tmpdir="$PWD")
annotatedTopHits <- fread(params$annotated_tophits_filename, select = c("ID", "TEST", "GENE_NAME"), header=TRUE, tmpdir="$PWD")

#merge gene names of tophits to regenieTable
annotatedTopHits <- annotatedTopHits %>%  
    mutate(ID = as.character(ID),
           TEST = as.character(TEST))
regenieTable <- regenieTable %>%  
    mutate(ID = as.character(ID),
           TEST = as.character(TEST))

regenieTable_annotated <- left_join(regenieTable, annotatedTopHits, by = c("ID",
                                                                           "TEST"))
#subset by test variable

split_regenieTable_annotated <- split(regenieTable_annotated, regenieTable_annotated$TEST)
rm(regenieTable_annotated)

#generate manhattan and qq plots by test
for (test in 1:length(split_regenieTable_annotated)) {
  test_name <- unique(split_regenieTable_annotated[[test]]$TEST)
  cat('\n\n')
  cat('### Test: ', test_name)
  cat('\n\n')
  #manhattan plot
  cat('\n\n')
  cat('**Manhattan Plot**')
  cat('\n\n')
  df_prep <- manhattan_prep(split_regenieTable_annotated[[test]])
  plot <- manhattan_plot(df_prep)
  if(!params$manhattan_annotation_enabled) {
  plot
  } else {
    plot <- plot +
      ggrepel::geom_text_repel(data = df_prep %>% group_by(GENE_NAME) %>%
                                 top_n(1, LOG10P) %>%
                                 slice(1) %>%
                                 filter(LOG10P > params$annotation_min_log10p),
                               aes(label=GENE_NAME), color = 1, fontface = "italic")
    }
  print(plot)
  #qq plot
  cat('\n\n')
  cat('**QQ Plot**')
  cat('\n\n')
  qqPrepare <- qqPlotPrepare(pvalues=split_regenieTable_annotated[[test]]$LOG10P,ismlog10 = T)
  qqPlotFast(qqPrepare)
}
```


## Top Loci

Filtered for a minimun -log<sub>10</sub>(p-value) = `r params$annotation_min_log10p` (no loci are depicted if there are no results above this threshold).

```{r, echo=FALSE,include = FALSE}
DT::datatable(matrix())
```

```{r, results='asis'}

annotatedTopHits <- fread(params$annotated_tophits_filename, select = c("CHROM", "GENPOS", "TEST", "LOG10P", "ID", "GENE_NAME", "GENE_CHROMOSOME", "GENE_START", "GENE_END", "DISTANCE"), header=TRUE, tmpdir="$PWD")

split_annotatedTopHits <- split(annotatedTopHits, annotatedTopHits$TEST)

for (test in 1:length(split_annotatedTopHits)) {
  test_name <- unique(split_annotatedTopHits[[test]]$TEST)
  cat('\n\n')
  cat('### Test: ', test_name)
  cat('\n\n')
  #Top loci
  annotatedTopHits_1test <- split_annotatedTopHits[[test]]
  
  topHitsGroupedByGene <-
  annotatedTopHits_1test %>%
  group_by(GENE_NAME, GENE_CHROMOSOME, GENE_START, GENE_END) %>%
  summarise(GENE_LOG10P = max(LOG10P), VARIANTS = n()) %>%
  arrange(desc(GENE_LOG10P))
  
  cat(knitr::knit_print(datatable(topHitsGroupedByGene, width = 
                                    "100%")))
  cat("\n\n")
}

```

## Validation and Logs

```{r child=params$rmd_valdiation_logs}

```


---

<small>
This report has been created with **[nf-gwas `r params$version`](https://github.com/genepi/nf-gwas)**, a nextflow pipeline developed by [Sebastian Schönherr](mailto:sebastian.schoenherr@i-med.ac.at), [Lukas Forer](mailto:lukas.forer@i-med.ac.at) and [Johanna Schachtl-Riess](mailto:johanna.schachtl-riess@i-med.ac.at), Institute of Genetic Epidemiology, Medical University of Innsbruck.
</small>
