---
title: "Report: `r params$project`"
output:
  rmdformats::robobook:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    highlight: tango
params:
  project: test-gwas
  date: 2021-08-09
  version: v0.0.1
  regenie_merged: ../results/test-gwas/regenie_results/test-gwas.Y1.regenie.all.gz
  regenie_filename: test-gwas.Y1.regenie.all.gz
  phenotype: lpa_cleaned_normal
  covariates:
  phenotype_file: /home/seb/Desktop/lpa_phenotype.txt
  regenie_step1_log: ../results/test-gwas/regenie_logs/test-gwas.step1.log
  regenie_step2_log: ../results/test-gwas/regenie_logs/test-gwas.step2.log
---

```{r setup, include=FALSE}
library(qqman)
library(data.table)
library(R.utils)
library(ggplot2)
library(skimr)
library(kableExtra)
library(ramwas)
```

## Project Summary

| Parameter     | Value                                      |
| ------------- |--------------------------------------------|
| Project       | `r params$project`                         |
| Pipeline Version       | `r params$version`                |
| Date       | `r params$date`                         |
| Phenotype File      | `r params$phenotype_file`            |
| Phenotype       | `r params$phenotype`                     |
| Covariates       | `r params$covariates`                     |
| Regenie Output      | `r params$regenie_filename`             |

## Regenie Log

### Step 1

```{r, echo=FALSE, results='asis'}

if(file.info(params$regenie_step1_log)$size != 0) {
step1_log <- read.table(params$regenie_step1_log,sep ="\t", header = TRUE, dec =".")
kable(step1_log)
} else {
    cat("*No log file available, since regenie step 1 is skipped.*")
}
```

### Step 2

```{r, echo=FALSE}
step2_log <- read.table(params$regenie_step2_log,
                      sep ="\t", header = TRUE, dec =".")
kable(step2_log)
```

## Phenotype Statistics

### Overview

```{r, echo=FALSE}
phenotypeTable <- read.table(text = gsub(" ", "\t", readLines(params$phenotype_file)), header = TRUE)
kable(skim(phenotypeTable) %>%
  dplyr::filter(skim_variable == params$phenotype)) %>%
  kableExtra::scroll_box(width = "100%")
```

### Histogram
```{r, echo=FALSE}
params$phenotype
phenotypeTable
aes_string(x=params$phenotype)
ggplot(phenotypeTable, aes_string(x=params$phenotype)) +
geom_histogram(color="black", fill="white",bins=30)
```

## Manhattan Plot

```{r, echo=FALSE}
regenieTable <- fread(params$regenie_merged, select = c("CHROM","GENPOS","LOG10P","ID"), header=TRUE)
pheno <- params$phenotype
regenieTable
manhattan(regenieTable, chr="CHROM", bp="GENPOS", p="LOG10P", snp="ID", logp= FALSE, cex = 0.3, col = c("blue4", "orange3"))
```

## QQ Plot

```{r, echo=FALSE}
qqPrepare <- qqPlotPrepare(pvalues=regenieTable$LOG10P,ismlog10 = T)
qqPlotFast(qqPrepare)
```

## About 
This report has been created with a GWAS Nextflow Pipeline (https://github.com/genepi/gwas-regenie) (**`r params$version`**) by Lukas Forer (lukas.forer@i-med.ac.at) and Sebastian Schönherr (sebatian.schoenherr@i-med.ac.at), Institute of Genetic Epidemiology, Medical University of Innsbruck.
